---
title: "Credit Card Defaults - Metrics Comparison"
author: "Miles Rollison"
date: "3/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
library(lemon)
knit_print.data.frame = lemon_print
options(scipen = 8)
options(digits = 4)

set.seed(8675309)
```

```{r python setup, include=F}
# Sys.setenv(RETICULATE_PYTHON = "python/bin/python")
# reticulate::py_config()
```

```{r packages, message=F}
library(tidyverse)
library(magrittr)
library(RSkittleBrewer)
```

```{r data import, message=F}
credit = read_csv('CCDefaults_Yeh_Lien.csv')
```

```{r data cleaning, message=F}
colnames(credit) %<>% tolower() %>% make.names()
credit %<>%
  rename(default_oct = default.payment.next.month) %>% 
  mutate_at(vars(sex, marriage), function(x){x - 1})
#NOTE: bill_amt1 = september; pay_amt1 = september
```

```{r feature construction, message=F}
credit %<>%
  mutate(balance_1 = bill_amt1-pay_amt1,
       balance_2 = bill_amt2-pay_amt2,
       balance_3 = bill_amt3-pay_amt3,
       balance_4 = bill_amt4-pay_amt4,
       balance_5 = bill_amt5-pay_amt5,
       balance_6 = bill_amt6-pay_amt6)

credit %<>%
  rowwise() %>%
  mutate(avg_pay = mean(c(pay_amt6, pay_amt5, pay_amt4, pay_amt3, pay_amt2, pay_amt1)),
         median_pay = median(c(pay_amt6, pay_amt5, pay_amt4, pay_amt3, pay_amt2, pay_amt1)),
         avg_bill = mean(c(bill_amt6, bill_amt5, bill_amt4, bill_amt3, bill_amt2, bill_amt1)),
         median_bill = median(c(bill_amt6, bill_amt5, bill_amt4, bill_amt3, bill_amt2, bill_amt1)),
         avg_3_bill = mean(c(bill_amt3, bill_amt2, bill_amt1)),
         avg_3_pay = mean(c(pay_amt3, pay_amt2, pay_amt1)),
         avg_3_balance = mean(c(balance_3, balance_2, balance_1)))
```

```{r quantiles}
quantiles = credit %>% 
  select(avg_3_balance) %>%
  filter(avg_3_balance > 0) %$% 
  quantile(avg_3_balance, probs = c(0.2, 0.4, 0.6, 0.8, .95))

# credit %<>% 
#   mutate(percentile_bin = case_when(avg_3_balance == 0 ~ 0,
#                                       avg_3_balance < quantiles[1] ~ 20,
#                                       avg_3_balance >= quantiles[1] & avg_3_balance < quantiles[2] ~ 40,
#                                       avg_3_balance >= quantiles[2] & avg_3_balance < quantiles[3] ~ 60,
#                                       avg_3_balance >= quantiles[3] & avg_3_balance < quantiles[4] ~ 80,
#                                       avg_3_balance >= quantiles[4] & avg_3_balance < quantiles[5]~ 95,
#                                       avg_3_balance >= quantiles[5] ~ 100,
#                                       TRUE ~ -1))

credit %<>%
  mutate(percentile_bin = case_when(avg_3_balance <= 0 ~ "0",
                                      avg_3_balance > 0 & avg_3_balance < quantiles[1] ~ "1-19",
                                      avg_3_balance >= quantiles[1] & avg_3_balance < quantiles[2] ~ "20-39",
                                      avg_3_balance >= quantiles[2] & avg_3_balance < quantiles[3] ~ "40-59",
                                      avg_3_balance >= quantiles[3] & avg_3_balance < quantiles[4] ~ "60-79",
                                      avg_3_balance >= quantiles[4] & avg_3_balance < quantiles[5]~ "79-94",
                                      avg_3_balance >= quantiles[5] ~ "95-100",
                                      TRUE ~ "-1"))


# credit %>% group_by(percentile_bin) %>% summarise(median(avg_3_balance),
#                                                   n())

#credit[which(credit$percentile_bin == "-1" & credit$default_oct == 1),]
```


```{r retype}
credit %<>% mutate_at(vars(id, sex, education, marriage, age,
                          pay_0, pay_2, pay_3, pay_4, pay_5, pay_6,
                          default_oct), as.integer)
```

```{r data export}
arrow::write_parquet(credit, "CCDefaults.parquet")
```


## Data
The dataset used in this analysis is the [Default of Credit Card Clients](https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients) dataset provided by Yeh, I. C., & Lien, C. H. (2009) via the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/index.php). 

Attribute Information:

This research employed a binary variable, default payment (Yes = 1, No = 0), as the response variable. This study reviewed the literature and used the following 23 variables as explanatory variables:
X1: Amount of the given credit (NT dollar): it includes both the individual consumer credit and his/her family (supplementary) credit.
X2: Gender (1 = male; 2 = female).
X3: Education (1 = graduate school; 2 = university; 3 = high school; 4 = others).
X4: Marital status (1 = married; 2 = single; 3 = others).
X5: Age (year).
X6 - X11: History of past payment. We tracked the past monthly payment records (from April to September, 2005) as follows: X6 = the repayment status in September, 2005; X7 = the repayment status in August, 2005; . . .;X11 = the repayment status in April, 2005. The measurement scale for the repayment status is: -1 = pay duly; 1 = payment delay for one month; 2 = payment delay for two months; . . .; 8 = payment delay for eight months; 9 = payment delay for nine months and above.
X12-X17: Amount of bill statement (NT dollar). X12 = amount of bill statement in September, 2005; X13 = amount of bill statement in August, 2005; . . .; X17 = amount of bill statement in April, 2005.
X18-X23: Amount of previous payment (NT dollar). X18 = amount paid in September, 2005; X19 = amount paid in August, 2005; . . .;X23 = amount paid in April, 2005. 

## References

Yeh, I. C., & Lien, C. H. (2009). The comparisons of data mining techniques for the predictive accuracy of probability of default of credit card clients. Expert Systems with Applications, 36(2), 2473-2480.